---
title: "grandslamR - a brief tutorial"
author: "Peter Tea"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{grandslamR - a brief tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is a brief tutorial on loading and plotting the [courtvision tracking data](google.com). Please note that the `grandslamR` package is work in progress; if you have any suggestions or bugs to report, please let me know!

## Package Installation

To install the package, 
```{r, installation, echo = FALSE, message=FALSE}
#install.packages('grandslamR')
#library(grandslamR)
library(ggplot2)
library(dplyr)
devtools::load_all()
```

## Loading Data

There are two options from which you can load in data:
 - By a single match (using `match_id`)
 - By a single player (using `player_id`)
 
### Loading Data by Match

Loading by a single match requires you to input a valid `match_id`. All valid match IDs can be found as follows:

```{r, load_available_matches}
knitr::kable(head(grandslamR::load_available_matches()))
```

Note that this data source includes a few `doubles matches` -- although the majority of available data are singles matches. Let's take a closer look at the 2019 Roland Garros final between Thiem and Nadal. To load in data from this match, we will need to know it's `match_id`.

```{r, thiem_nadal_example}
thiem_nadal <- load_complete_match_data(match_id = 'atp_roland_garros_2019_SM001')
knitr::kable(head(thiem_nadal))
```

For a better understanding on what each column represents, please check the data dictionary.

### Loading Data by Player Name

```{r, shapovalov_example}
knitr::kable(head(load_complete_player_data(player_name = 'D.SHAPOVALOV')))
```


## Plotting Data

The `grandslamR` package also supports plotting the spatio-temporal ball tracking data. For example, the following plot illustrates the ball bounce locations on a single rally point (1st set, 7th game, 12th point on 1st serve).

```{r, thiem_nadal_plot, fig.width = 8, fig.height = 5}
thiem_nadal %>%
  dplyr::filter(point_ID == '1_7_12_1') %>%
  dplyr::filter(position == 'bounce') %>%
ggplot2::ggplot(aes(x = x,
           y = y)) +
  gg_fulltenniscourt() +
  ggplot2::geom_point(alpha = 0.75, fill = 'green', shape = 21,
             size = 2.5) +
  ggplot2::geom_path(color = 'white', linetype = 2)
```


### To Do List

Here are a list of items that will hopefully be added in this R package's development.

 - Add in Point Importance Feature  
 
 - Add in processing steps to plot all points on half of a tennis court
 
 - Add in functions to plot shot heatmaps
 
 - Add in functions to interpolate between ball locations
